---
title: "home_work2"
output: pdf_document
---

task1:

All information in summary:

```{r}
data = read.csv("data.csv", sep = ",")
summary(data)
colnames(data)
```

As we know age is one of the most important factors in hockey, a lot of other characteristics depends on it. Also the hockey experts usually compare the average age of teams and then make some assumptions about how the teams play.

histogram:

```{r, echo=FALSE}
hist(data$Age, 8)

```

boxplot:

```{r}
boxplot(data$Age, main = 'Boxplot of data$Age', horizontal = TRUE)

```

mean:

```{r}
mean(data$Age)
```
median:

```{r}
median(data$Age)
```

mode:

```{r}

findMode <- function(arg){
data.t <- table(arg)
return(sort(unique(arg))[which.max(data.t)])
}
findMode(data$Age)
```

Hockey clubs select players since 18 years and track them throughout their careers. But if a player has not proved himself up to a certain age, then club parted with him. The mode is 23 years, it shows that basically in this age clubs give players a last chance to show themselves. Later on most of them goes to the lower league. Therefore, there is a so large number of 23-year-old players in the league.
A median more than mode just because the maximum players' age is 42 years, and in addition to the 23-year-old players core of the team for the most part consits of the older players.

task2:

statistical mean:

```{r}
mean(data$Age)
n <- nrow(data)
m <- mean(data$Age)
s <- sd(data$Age)
conf.int <- c(m-1.965*s/sqrt(n), m+1.965*s/sqrt(n))
print(conf.int)
```

bootstrap:

```{r}
means <- c()
for(i in 1:5000)
  {
  means[i] <- mean(sample(data$Age, 50, replace = TRUE))
  }
```

new histogram:

```{r}
hist(means)
```

Because the histogram is similar to a normal distribution, we apply the technique of pivotal bootstrap.

```{r}
n <- 5000
m <- mean(means)
s <- sd(means)
pivotal.conf.int <- c(m-1.965*s/sqrt(n), m+1.965*s/sqrt(n))
print(pivotal.conf.int)
```

non-pivotal bootstrap:

```{r}
sorted.means  <- sort(means)
nonpivotal.conf.int <- c(sorted.means[0.025*5000], sorted.means[5000-0.025*5000])
print(nonpivotal.conf.int)
```

Statistical confidence interval [25.93271 26.54336]

task3:

pivotal bootstrap for median:

```{r}
medians <- c()
for(i in 1:5000)
  {
  medians[i] <- median(sample(data$Age, 50, replace = TRUE))
  }
hist(medians)
```

Because the histogram is not similar to a normal distribution, we can't apply the technique of pivotal bootstrap here.

Non-pivotal bootstrap for median:

```{r}
sorted.medians  <- sort(medians)
nonpivotal.conf.int <- c(sorted.medians[0.025*5000], sorted.medians[5000-0.025*5000])
print(nonpivotal.conf.int)
```

pivotal bootstrap for mode:

```{r}
modes <- c()
for(i in 1:5000)
  {
  modes[i] <- findMode(sample(data$Age, 50, replace = TRUE))
  }
hist(modes)
```

Because the histogram is not similar to a normal distribution, we can't apply the technique of pivotal bootstrap here.

Non-pivotal bootstrap for mode:

```{r}
sorted.modes  <- sort(modes)
nonpivotal.conf.int <- c(sorted.modes[0.025*5000], sorted.modes[5000-0.025*5000])
print(nonpivotal.conf.int)
```

task 4:

Devide players into two groups. Name players who has a salary higher than the median of all salaries in the league high-salary players, and who has a salary lower than the median - low-salary players. Consider their ages.

```{r}
high.salary.players  <- data[data$Salary >= median(data$Salary),]
low.salary.players  <- data[data$Salary < median(data$Salary),]
print (mean(high.salary.players$Age))
print (mean(low.salary.players$Age))
```


bootstrap for high.salary.means:

```{r}
high.salary.means <- c()
for(i in 1:5000)
  {
  high.salary.means[i] <- mean(sample(high.salary.players$Age, 50, replace = TRUE))
  }
hist(high.salary.means)
```

Because the histogram is similar to a normal distribution, we apply the technique of pivotal bootstrap.

```{r}
n <- 5000
m <- mean(high.salary.means)
s <- sd(high.salary.means)
pivotal.conf.int <- c(m-1.965*s/sqrt(n), m+1.965*s/sqrt(n))
print(pivotal.conf.int)
```

Non-pivotal bootstrap for high.salary.means:

```{r}
sorted.high.salary.means  <- sort(high.salary.means)
nonpivotal.conf.int <- c(sorted.high.salary.means[0.025*5000],
                         sorted.high.salary.means[5000-0.025*5000])
print(nonpivotal.conf.int)
```

bootstrap for low.salary.means:

```{r}
low.salary.means <- c()
for(i in 1:5000)
  {
  low.salary.means[i] <- mean(sample(low.salary.players$Age, 50, replace = TRUE))
  }
hist(low.salary.means)
```

Because the histogram is similar to a normal distribution, we apply the technique of pivotal bootstrap.

```{r}
n <- 5000
m <- mean(low.salary.means)
s <- sd(low.salary.means)
pivotal.conf.int <- c(m-1.965*s/sqrt(n), m+1.965*s/sqrt(n))
print(pivotal.conf.int)
```

Non-pivotal bootstrap for high.salary.means:

```{r}
sorted.low.salary.means  <- sort(low.salary.means)
nonpivotal.conf.int <- c(sorted.low.salary.means[0.025*5000],
                         sorted.low.salary.means[5000-0.025*5000])
print(nonpivotal.conf.int)
```

Confidence intervals differs considerably. It is clear, that high-salary players are older than low-salary players on the average, because older players has more experience and, hense, are more valuable.

